---
title: "Spatial-Temporal Regression Models"
output:
  rmarkdown::html_vignette:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
vignette: >
  %\VignetteIndexEntry{Spatial-Temporal Regression Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
  - \def\T{{ \scriptstyle \top }}
  - \newcommand{\thetasp}{{\theta_{\text{sp}}}}
  - \newcommand{\GP}{\mathrm{GP}}
  - \newcommand{\N}{\mathrm{N}}
  - \newcommand{\EF}{\mathrm{EF}}
  - \newcommand{\Norm}{\mathrm{N}}
  - \newcommand{\GCMc}{\mathrm{GCM}_c}
  - \newcommand{\calL}{\mathcal{L}}
  - \newcommand{\IG}{\mathrm{IG}}
  - \newcommand{\IW}{\mathrm{IW}}
  - \newcommand{\given}{\mid}
---

In this article, we discuss the following functions -

- `stvcGLMexact()`
- `stvcGLMstack()`
- `recoverGLMscale()`

These functions can be used to fit non-Gaussian spatial-temporal point-referenced data.

```{r}
set.seed(1729)
```


## Bayesian non-Gaussian spatially-temporally varying coefficient models

We illustrate the spatially-temporally varying coefficient model using the synthetic spatial-temporal Poisson count data.

We first load the data `sim_stvcPoisson` which consists of data at 500 spatial-temporal locations. We use the first 100 locations for the following analysis.
```{r}
library(spStack)
data("sim_stvcPoisson")
n_train <- 100
dat <- sim_stvcPoisson[1:n_train, ]
```

The dataset consists of one covariate `x1`, response variable `y`, spatial locations given by `s1` and `s2`, a temporal coordinate `t_coords`, and the true spatially-temporally varying coefficients `z1_true` and `z2_true` associated with an intercept and `x1`, respectively. We elaborate below. 
```{r}
head(dat)
```

### Formula for varying coefficients model
We define the spatially-temporally varying coefficients model using a `formula`, similar to that in the widely used `lm()` function in the `stats` package. Suppose $\ell = (s, t)$ refers to a space-time ccoordinate. See "Technical Overview for more details". Then, given `family = "poisson"`, the formula `y ~ x1 + (x1)` corresponds to the spatial-temporal generalized linear model 
$$y(\ell) \sim \mathsf{Poisson}(\lambda(\ell)), \quad \log \lambda(\ell) = \beta_0 + \beta_1 x_1(\ell) + z_1(\ell) + x_1(\ell) z_2(\ell)\;,$$ where the `y` corresponds to the response variable $y(\ell)$, which is regressed on the predictor `x1` given by $x_1(\ell)$. The model variables specified outside the parentheses corresponds to predictors with fixed effects, and the model inside the parentheses correspond to variables with spatial-temporal varying coefficient. The intercept is automatically considered within both the fixed and varying coefficient components of the model, and hence `y ~ x1 + (x1)` is functionally equivalent to `y ~ 1 + x1 + (1 + x1)`. The spatially-temporally varying coefficients $z(\ell) = (z_1(\ell), z_2(\ell))^{\T}$ is multivariate Gaussian process, and we pursue the following specifications for $z(\ell)$ - independent process, independent process with shared parameters, and a multivariate process. For now, we only support the `cor.fn="gneiting-decay"` covariogram. See "Technical Overview" for more details.

To implement a model, with just a spatial-temporal random effect, one may specify the formula `y ~ x1 + (1)` which corresponds to the model
$$y(\ell) \sim \mathsf{Poisson}(\lambda(\ell)), \quad \log \lambda(\ell) = \beta_0 + \beta_1 x_1(\ell) + z_1(\ell)\;.$$

### Using fixed hyperparameters
We use the function `stvcGLMexact()` to fit spatially-temporally varying coefficient generalized linear models. In the following code snippets, we demonstrate the uasge of the argument `process.Type` to implement different variations of spatial-temporal process specifications for the varying coefficients.

#### Independent processes
In this case, since there are two independent processes $z_1(\ell)$ and $z_2(\ell)$ the candidate values of the spatial-temporal process parameters `sptParams` is a list with tags `phi_s` and `phi_t`, with each tag being of length 2. Here, the scale parameter $\sigma = (\sigma^2_{z1}, \sigma^2_{z2})^{\T}$ has dimension 2.

```{r}
mod1 <- stvcGLMexact(y ~ x1 + (x1), data = dat, family = "poisson",
                     sp_coords = as.matrix(dat[, c("s1", "s2")]),
                     time_coords = as.matrix(dat[, "t_coords"]),
                     cor.fn = "gneiting-decay",
                     process.type = "independent",
                     priors = list(nu.beta = 5, nu.z = 5),
                     sptParams = list(phi_s = c(1, 2), phi_t = c(1, 2)),
                     verbose = FALSE, n.samples = 500)
```

Posterior samples of the scale parameters can be recovered by running `recoverGLMscale()` on `mod1`.
```{r}
mod1 <- recoverGLMscale(mod1)
```

We visualize the posterior distributions of the scale parameters as follows.

```{r fig.align='center', fig.height=3.5, fig.width=7, fig.alt="Posterior distributions of the scale parameters."}
post_scale_df <- data.frame(value = sqrt(c(mod1$samples$z.scale[1, ], mod1$samples$z.scale[2, ])),
                            group = factor(rep(c("sigma.z1", "sigma.z2"), 
                                    each = length(mod1$samples$z.scale[1, ]))))
library(ggplot2)
ggplot(post_scale_df, aes(x = value)) +
  geom_density(fill = "lightblue", alpha = 0.6) +
  facet_wrap(~ group, scales = "free") + labs(x = "", y = "Density") +
  theme_bw() + theme(panel.background = element_blank(), 
                     panel.grid = element_blank(), aspect.ratio = 1)
```

#### Independent shared processes
In this case, the processes $z_1(\ell)$ and $z_2(\ell)$ are independent but share a common covariance matrix. Hence, `sptParams` is a list with tags `phi_s` and `phi_t`, with each tag being of length 1. Here, the scale parameter $\sigma = \sigma_z^2$ is 1-dimensional.

```{r}
mod2 <- stvcGLMexact(y ~ x1 + (x1), data = dat, family = "poisson",
                     sp_coords = as.matrix(dat[, c("s1", "s2")]),
                     time_coords = as.matrix(dat[, "t_coords"]),
                     cor.fn = "gneiting-decay",
                     process.type = "independent.shared",
                     priors = list(nu.beta = 5, nu.z = 5),
                     sptParams = list(phi_s = 1, phi_t = 1),
                     verbose = FALSE, n.samples = 500)
```

Posterior samples of the scale parameters can be recovered by running `recoverGLMscale()` on `mod2`.
```{r}
mod2 <- recoverGLMscale(mod2)
```

We visualize the posterior distributions of the scale parameters as follows.

```{r fig.align='center', fig.height=3.5, fig.width=7, fig.alt="Posterior distributions of the scale parameters."}
post_scale_df <- data.frame(value = sqrt(mod2$samples$z.scale), 
                            group = factor(rep(c("sigma.z"), 
                                               each = length(mod2$samples$z.scale))))
ggplot(post_scale_df, aes(x = value)) +
  geom_density(fill = "lightblue", alpha = 0.6) +
  facet_wrap(~ group, scales = "free") + labs(x = "", y = "Density") +
  theme_bw() + theme(panel.background = element_blank(), 
                     panel.grid = element_blank(), aspect.ratio = 1)
```

#### Multivariate processes
In this case, $z(\ell) = (z_1(\ell), z_2(\ell))^{\T}$ is a 2-dimensional Gaussian process with covariance matrix $\Sigma$. Further, we put an inverse-Wishart prior on $\Sigma$, which can be specified through the `priors` argument. If not supplied, uses the default $\IW (\nu_z + 2r, I_r)$, where $r = 2$ is the dimension of the multivariate process. Here, `sptParams` is a list with tags `phi_s` and `phi_t`, with each tag being of length 1, and the scale parameter $\sigma = \Sigma$ is an $2 \times 2$ matrix.

```{r}
mod3 <- stvcGLMexact(y ~ x1 + (x1), data = dat, family = "poisson",
                     sp_coords = as.matrix(dat[, c("s1", "s2")]),
                     time_coords = as.matrix(dat[, "t_coords"]),
                     cor.fn = "gneiting-decay",
                     process.type = "multivariate",
                     priors = list(nu.beta = 5, nu.z = 5),
                     sptParams = list(phi_s = 1, phi_t = 1),
                     verbose = FALSE, n.samples = 500)
```

Posterior samples of the scale parameters can be recovered by running `recoverGLMscale()` on `mod3`.
```{r}
mod3 <- recoverGLMscale(mod3)
```

We visualize the posterior distribution of the scale matrix $\Sigma$ as follows.

```{r fig.align='center', fig.height=5, fig.width=4, fig.cap="Posterior distributions of elements of the scale matrix."}
post_scale_z <- mod3$samples$z.scale

r <- sqrt(dim(post_scale_z)[1])
# Function to get (i,j) index from row number (column-major)
get_indices <- function(k, r) {
  j <- ((k - 1) %/% r) + 1
  i <- ((k - 1) %% r) + 1
  c(i, j)
}

# Generate plots into a matrix
plot_matrix <- matrix(vector("list", r * r), nrow = r, ncol = r)
for (k in 1:(r^2)) {
  ij <- get_indices(k, r)
  i <- ij[1]
  j <- ij[2]
  
  if (i >= j) {
    df <- data.frame(value = post_scale_z[k, ])
    p <- ggplot(df, aes(x = value)) +
      geom_density(fill = "lightblue", alpha = 0.7) +
      theme_bw(base_size = 9) +
      labs(title = bquote(Sigma[.(i) * .(j)])) +
      theme(axis.title = element_blank(), axis.text = element_text(size = 6),
        plot.title = element_text(size = 9, hjust = 0.5),
        panel.grid = element_blank(), aspect.ratio = 1)
  } else {
    p <- ggplot() + theme_void()
  }
  
  plot_matrix[j, i] <- list(p)
}

library(patchwork)
# Assemble with patchwork
final_plot <- wrap_plots(plot_matrix, nrow = r)
final_plot
```

### Using predictive stacking

For implementing predictive stacking for spatially-temporally varying models, we offer a helper function `candidateModels()` to create a collection of candidate models. The grid of candidate values can be combined either using a Cartesian product or a simple element-by-element combination. We demonstrate stacking based on the multivariate spatial-temporal process model.

**Step 1.** Create candidate models.
```{r}
mod.list <- candidateModels(list(
  phi_s = list(1, 2, 3),
  phi_t = list(1, 2, 4),
  boundary = c(0.5, 0.75)), "cartesian")
```

**Step 2.** Run `stvcGLMstack()`.
```{r}
mod1 <- stvcGLMstack(y ~ x1 + (x1), data = dat, family = "poisson",
                     sp_coords = as.matrix(dat[, c("s1", "s2")]),
                     time_coords = as.matrix(dat[, "t_coords"]),
                     cor.fn = "gneiting-decay",
                     process.type = "multivariate",
                     priors = list(nu.beta = 5, nu.z = 5),
                     candidate.models = mod.list,
                     loopd.controls = list(method = "CV", CV.K = 10, nMC = 500),
                     n.samples = 1000)
```

**Step 3.** Recover posterior samples of the scale parameters.
```{r}
mod1 <- recoverGLMscale(mod1)
```

**Step 4.** Sample from the stacked posterior distribution.
```{r}
post_samps <- stackedSampler(mod1)
```

Now, we analyze the posterior distribution of the latent process as obtained from the stacked posterior.
```{r fig.align='center', fig.height=3.5, fig.width=7}
post_z <- post_samps$z

post_z1_summ <- t(apply(post_z[1:n_train,], 1,
                        function(x) quantile(x, c(0.025, 0.5, 0.975))))
post_z2_summ <- t(apply(post_z[n_train + 1:n_train,], 1,
                        function(x) quantile(x, c(0.025, 0.5, 0.975))))

z1_combn <- data.frame(z = dat$z1_true, zL = post_z1_summ[, 1],
                       zM = post_z1_summ[, 2], zU = post_z1_summ[, 3])
z2_combn <- data.frame(z = dat$z2_true, zL = post_z2_summ[, 1],
                       zM = post_z2_summ[, 2], zU = post_z2_summ[, 3])

plot_z1_summ <- ggplot(data = z1_combn, aes(x = z)) +
  geom_errorbar(aes(ymin = zL, ymax = zU), alpha = 0.5, color = "skyblue") +
  geom_point(aes(y = zM), size = 0.5, color = "darkblue", alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "solid") +
  xlab("True z1") + ylab("Posterior of z1") + theme_bw() +
  theme(panel.grid = element_blank(), aspect.ratio = 1)

plot_z2_summ <- ggplot(data = z2_combn, aes(x = z)) +
  geom_errorbar(aes(ymin = zL, ymax = zU), alpha = 0.5, color = "skyblue") +
  geom_point(aes(y = zM), size = 0.5, color = "darkblue", alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "solid") +
  xlab("True z2") + ylab("Posterior of z2") + theme_bw() +
  theme(panel.grid = element_blank(), aspect.ratio = 1)

ggpubr::ggarrange(plot_z1_summ, plot_z2_summ)
```

Next, we analyze the posterior distribution of the scale matrix that models the inter-process dependence structure.

```{r fig.align='center', fig.height=5, fig.width=4, fig.cap="Stacked posterior distribution of the elements of the inter-process covariance matrix."}
post_scale_z <- post_samps$z.scale
r <- sqrt(dim(post_scale_z)[1])
# Generate plots into a matrix
plot_matrix <- matrix(vector("list", r * r), nrow = r, ncol = r)
for (k in 1:(r^2)) {
  ij <- get_indices(k, r)
  i <- ij[1]
  j <- ij[2]
  
  if (i >= j) {
    df <- data.frame(value = post_scale_z[k, ])
    p <- ggplot(df, aes(x = value)) +
      geom_density(fill = "lightblue", alpha = 0.7) +
      theme_bw(base_size = 9) +
      labs(title = bquote(Sigma[.(i) * .(j)])) +
      theme(axis.title = element_blank(), axis.text = element_text(size = 6),
        plot.title = element_text(size = 9, hjust = 0.5),
        panel.grid = element_blank(), aspect.ratio = 1)
  } else {
    p <- ggplot() + theme_void()
  }
  
  plot_matrix[j, i] <- list(p)
}

# Assemble with patchwork
final_plot <- wrap_plots(plot_matrix, nrow = r)
final_plot
```

